Covariate analysis of CMC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (ACC and DLPFC with CQN normlaisation)

### 2017-07-28 code/Rmd/CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_SVA.Rmd
Minor changes

knit2synapse::knitToFolderEntity(file = "./CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_SVA.Rmd",
                                 parentId = "syn9872607",
                                 entityName = 'ACC and DLPFC - STAR - FEATURE COUNTS - ENSEMBL v75 - CQN - SVA')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = "syn9872607";
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of STAR aligned feature counts with Ensembl v75 with CQN normalisation (ACC and DLPFC)';

@@

# Compute actual variance of all principal components
tmp = svd(RESIDUAL.GENE_EXPRESSION %>% changeNAtoZero)
actual.var = tmp$d^2/sum(tmp$d^2)

# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind, .packages = c('dplyr')) %dopar% {
  tmp.residual = t(apply(RESIDUAL.GENE_EXPRESSION %>% changeNAtoZero, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])

@@

# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$E %>% changeNAtoZero, MODEL1, MODEL0, n.sv = NUM.SV, B = 30, tol = 1e-20)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)

# Estimate voom weights for dispersion control
VOOM.SVA.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts %>% changeNAtoZero, 
                        design=cbind(MODEL1, SURR.VAR), plot=F)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

@@

# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts %>% changeNAtoZero, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Individual_ID, 
                        correlation = correlation$cor)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

@@

# Store enrichment results
write.table(enrichment.results, file = 'CMC_ACC_DLPFC_EnrichResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('CMC_ACC_DLPFC_EnrichResults.tsv', 
                  name = 'Enrichment Analysis Results (Dx.Tissue)', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichResults'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Residual Expression (for network analysis) |  `r nEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Analysis                        |  `r ENRICH_OBJ$properties$id`    |

>>>>

knit2synapse::knitToFolderEntity(file = "./CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_SVA.Rmd",
                                 parentId = "syn7501880",
                                 entityName = 'ACC and DLPFC - STAR - FEATURE COUNTS - ENSEMBL v75 - CQN - SVA')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = "syn7501880";
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of STAR aligned feature counts with Ensembl v75 with CQN normalisation (ACC and DLPFC)';

@@

# Compute actual variance of all principal components
tmp = svd(RESIDUAL.GENE_EXPRESSION)
actual.var = tmp$d^2/sum(tmp$d^2)

# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind, .packages = c('dplyr')) %dopar% {
  tmp.residual = t(apply(RESIDUAL.GENE_EXPRESSION, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])

@@

# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$E, MODEL1, MODEL0, n.sv = NUM.SV, B = 30, tol = 1e-20)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)

# Estimate voom weights for dispersion control
VOOM.SVA.WEIGHTS = voom(NEW.COUNTS, 
                        design=cbind(MODEL1, SURR.VAR), plot=F)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

@@

# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(NEW.COUNTS, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Individual_ID, 
                        correlation = correlation$cor)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

@@

# Store enrichment results
write.table(enrichment.results, file = 'CMC_ACC_DLPFC_EnrichResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('CMC_ACC_DLPFC_EnrichResults.tsv', 
                  name = 'Enrichment Analysis Results (Dx.Tissue)', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichResults'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Residual Expression (for eQTL analysis)    |  `r eEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Analysis                        |  `r ENRICH_OBJ$properties$id`    |
